### 2020-08-14 supplimentary/supplimentary_tables.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Supplimentary materials},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{float}
\usepackage{longtable}
\pagenumbering{gobble}
\usepackage{caption}
\captionsetup[table]{labelformat=empty}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Supplimentary materials}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\textbf{S1 Table. Summary of modifications to LESCA items.}

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-2}}
\centering
\begin{tabular}[t]{>{\raggedleft\arraybackslash}p{0.3cm}|>{\raggedright\arraybackslash}p{4.8cm}|>{\raggedright\arraybackslash}p{4.8cm}|>{\raggedright\arraybackslash}p{4.8cm}}
\hline
Q & Original & Students & Non-students\\
\hline
19 & Beginning a new school experience  (beginning college, transferring college etc) & Beginning a new school experience (beginning university, transferring university etc) & Beginning a new work experience\\
\hline
21 & Academic probation / ineligibility & \textit{No change} & Removed\\
\hline
22 & Being dismissed from dorm or other residence & Being dismissed from halls or other residence & Being asked to vacate house/home\\
\hline
27 & Financial problems concerning school & \textit{No change} & Financial problems\\
\hline
29 & Conflict with room-mate & \textit{No change} & Conflict within household\\
\hline
36 & Suspended from team for non-academic reasons & \textit{No change} & Removed\\
\hline
49 & Being absent from school (classes) because of participation in sport & Being absent from university (classes) because of participation in sport & Being absent from work because of participation in sport\\
\hline
61 & Major change in level of academic performance (doing better or worse) & \textit{No change} & Major change in level of performance at work (doing better or worse)\\
\hline
62 & Making career decisions (applying to graduate school,interviewing for jobs, etc) & Making career decisions (applying for Masters degree, interviewing for jobs, etc) & Making career decisions (applying for Masters degree, interviewing for jobs, etc)\\
\hline
\end{tabular}
\end{table}

\newpage

\textbf{S2 Table. All arc strengths greater than 0.3 included in the first network.}

\begin{longtable}[t]{l|l|r|r}
\caption{\label{tab:unnamed-chunk-3}}\\
\hline
from & to & strength & direction\\
\hline
\endfirsthead
\caption[]{ S2 Table continued}\\
\hline
from & to & strength & direction\\
\hline
\endhead
nlebase & nlelg\_1 & 1.00 & 1.00\\
\hline
nlebase & nlelg\_2 & 1.00 & 1.00\\
\hline
nlelg\_1 & injured\_1 & 1.00 & 1.00\\
\hline
nlelg\_2 & injured\_2 & 1.00 & 1.00\\
\hline
BIS\_1 & FFFS\_1 & 0.98 & 0.70\\
\hline
FFFS\_1 & BIS\_1 & 0.98 & 0.30\\
\hline
ind\_team & hours & 0.94 & 0.50\\
\hline
hours & ind\_team & 0.94 & 0.50\\
\hline
nlebase & RI\_1 & 0.84 & 1.00\\
\hline
ind\_team & nlebase & 0.78 & 1.00\\
\hline
BIS\_2 & FFFS\_2 & 0.74 & 0.71\\
\hline
FFFS\_2 & BIS\_2 & 0.74 & 0.29\\
\hline
gender & stiffness\_1 & 0.71 & 1.00\\
\hline
pi & stiffness\_1 & 0.57 & 1.00\\
\hline
nlelg\_1 & BIS\_1 & 0.55 & 0.97\\
\hline
BIS\_1 & nlelg\_1 & 0.55 & 0.03\\
\hline
FFFS\_1 & BIS\_2 & 0.52 & 1.00\\
\hline
clevel & balance\_1 & 0.47 & 1.00\\
\hline
injured\_2 & stiffness\_2 & 0.46 & 0.66\\
\hline
stiffness\_2 & injured\_2 & 0.46 & 0.34\\
\hline
injured\_1 & stiffness\_1 & 0.46 & 0.36\\
\hline
stiffness\_1 & injured\_1 & 0.46 & 0.64\\
\hline
gender & stiffness\_2 & 0.43 & 1.00\\
\hline
FFFS\_1 & injured\_2 & 0.42 & 1.00\\
\hline
hours & injured\_1 & 0.42 & 1.00\\
\hline
clevel & RI\_2 & 0.42 & 1.00\\
\hline
nlebase & hrv\_1 & 0.41 & 1.00\\
\hline
hours & hrv\_1 & 0.40 & 1.00\\
\hline
hrv\_1 & nlelg\_2 & 0.39 & 1.00\\
\hline
gender & FFFS\_1 & 0.37 & 1.00\\
\hline
injured\_2 & hrv\_2 & 0.37 & 0.89\\
\hline
nlelg\_2 & BIS\_2 & 0.37 & 0.96\\
\hline
BIS\_2 & nlelg\_2 & 0.37 & 0.04\\
\hline
hrv\_2 & injured\_2 & 0.37 & 0.11\\
\hline
pi & nlelg\_2 & 0.36 & 1.00\\
\hline
RI\_1 & BIS\_1 & 0.36 & 0.41\\
\hline
BIS\_1 & RI\_1 & 0.36 & 0.59\\
\hline
nlebase & RI\_2 & 0.36 & 1.00\\
\hline
injured\_1 & balance\_1 & 0.36 & 0.70\\
\hline
balance\_1 & injured\_1 & 0.36 & 0.30\\
\hline
injured\_2 & balance\_2 & 0.34 & 0.94\\
\hline
balance\_2 & injured\_2 & 0.34 & 0.06\\
\hline
RI\_1 & balance\_1 & 0.33 & 0.22\\
\hline
balance\_1 & RI\_1 & 0.33 & 0.78\\
\hline
gender & FFFS\_2 & 0.33 & 1.00\\
\hline
pi & clevel & 0.32 & 0.50\\
\hline
clevel & pi & 0.32 & 0.50\\
\hline
ind\_team & FFFS\_1 & 0.31 & 1.00\\
\hline
BIS\_1 & FFFS\_2 & 0.31 & 1.00\\
\hline
pi & ind\_team & 0.30 & 0.46\\
\hline
ind\_team & pi & 0.30 & 0.54\\
\hline
\end{longtable}

\clearpage
\newpage

\textbf{S3 Table. All arc strengths greater than 0.3 included in the second network.}

\begin{longtable}[t]{l|l|r|r}
\caption{\label{tab:unnamed-chunk-4}}\\
\hline
from & to & strength & direction\\
\hline
BIS & FFFS & 1.00 & 0.76\\
\hline
FFFS & BIS & 1.00 & 0.25\\
\hline
hours & injured & 0.72 & 1.00\\
\hline
injured & stiffness & 0.67 & 1.00\\
\hline
clevel & BIS & 0.60 & 1.00\\
\hline
hrv & FFFS & 0.58 & 0.69\\
\hline
FFFS & hrv & 0.58 & 0.31\\
\hline
gender & stiffness & 0.56 & 1.00\\
\hline
injured & nlec & 0.45 & 1.00\\
\hline
pi & balance & 0.43 & 1.00\\
\hline
pi & injured & 0.38 & 1.00\\
\hline
stiffness & BIS & 0.38 & 0.93\\
\hline
BIS & stiffness & 0.38 & 0.07\\
\hline
RI & BIS & 0.35 & 0.29\\
\hline
BIS & RI & 0.35 & 0.71\\
\hline
pi & hrv & 0.33 & 1.00\\
\hline
\end{longtable}

\newpage

\textbf{S4 Table. Probabilities of injury conditional on the variables in the Markov blanket for injured\_1.}

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-5}}
\centering
\begin{tabular}[t]{l|l|l|l|l|l}
\hline
prob & clevel & hours & nlelg\_1 & stiffness\_1 & balance\_1\\
\hline
0.53 & club\_university\_county & High & Low & High & High\\
\hline
0.46 & national\_international & High & Low & High & Low\\
\hline
0.44 & national\_international & High & Low & High & High\\
\hline
0.39 & club\_university\_county & Low & High & High & High\\
\hline
0.37 & club\_university\_county & High & Low & High & Low\\
\hline
0.32 & national\_international & Low & High & High & Low\\
\hline
0.32 & club\_university\_county & High & High & High & High\\
\hline
0.31 & club\_university\_county & High & High & Low & High\\
\hline
0.3 & national\_international & Low & High & High & High\\
\hline
0.29 & club\_university\_county & Low & High & Low & High\\
\hline
0.25 & national\_international & High & High & High & Low\\
\hline
0.25 & national\_international & High & High & Low & Low\\
\hline
0.25 & club\_university\_county & Low & High & High & Low\\
\hline
0.24 & national\_international & High & High & Low & High\\
\hline
0.24 & national\_international & High & High & High & High\\
\hline
0.24 & national\_international & Low & High & Low & Low\\
\hline
0.22 & national\_international & Low & High & Low & High\\
\hline
0.19 & club\_university\_county & High & High & Low & Low\\
\hline
0.19 & club\_university\_county & High & High & High & Low\\
\hline
0.18 & club\_university\_county & Low & High & Low & Low\\
\hline
0.17 & club\_university\_county & Low & Low & High & High\\
\hline
0.16 & club\_university\_county & High & Low & Low & High\\
\hline
0.13 & national\_international & Low & Low & High & Low\\
\hline
0.12 & national\_international & High & Low & Low & Low\\
\hline
0.12 & national\_international & Low & Low & High & High\\
\hline
0.12 & national\_international & High & Low & Low & High\\
\hline
0.09 & club\_university\_county & Low & Low & High & Low\\
\hline
0.09 & club\_university\_county & High & Low & Low & Low\\
\hline
0.07 & club\_university\_county & Low & Low & Low & High\\
\hline
0.06 & national\_international & Low & Low & Low & Low\\
\hline
0.05 & national\_international & Low & Low & Low & High\\
\hline
0.04 & club\_university\_county & Low & Low & Low & Low\\
\hline
\end{tabular}
\end{table}

\newpage

\textbf{S5 Table. Probabilities of injury conditional on the variables in the Markov blanket for injured\_2.}

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-6}}
\centering
\begin{tabular}[t]{r|l|l|l|l|l}
\hline
prob & FFFS\_1 & nlelg\_2 & stiffness\_2 & hrv\_2 & balance\_2\\
\hline
0.53 & Low & Low & High & Low & High\\
\hline
0.46 & Low & High & High & Low & High\\
\hline
0.41 & Low & Low & High & High & High\\
\hline
0.39 & Low & Low & High & Low & Low\\
\hline
0.35 & Low & High & High & High & High\\
\hline
0.33 & Low & High & High & Low & Low\\
\hline
0.32 & Low & Low & Low & Low & High\\
\hline
0.28 & Low & Low & High & High & Low\\
\hline
0.26 & Low & High & Low & Low & High\\
\hline
0.25 & High & Low & High & Low & High\\
\hline
0.23 & Low & High & High & High & Low\\
\hline
0.23 & Low & Low & Low & High & High\\
\hline
0.22 & High & High & High & Low & High\\
\hline
0.21 & Low & Low & Low & Low & Low\\
\hline
0.18 & Low & High & Low & High & High\\
\hline
0.17 & High & Low & High & High & High\\
\hline
0.17 & Low & High & Low & Low & Low\\
\hline
0.16 & High & Low & High & Low & Low\\
\hline
0.15 & High & High & High & High & High\\
\hline
0.14 & Low & Low & Low & High & Low\\
\hline
0.13 & High & High & High & Low & Low\\
\hline
0.13 & High & Low & Low & Low & High\\
\hline
0.11 & Low & High & Low & High & Low\\
\hline
0.11 & High & High & Low & Low & High\\
\hline
0.11 & High & Low & High & High & Low\\
\hline
0.09 & High & High & High & High & Low\\
\hline
0.08 & High & Low & Low & High & High\\
\hline
0.08 & High & Low & Low & Low & Low\\
\hline
0.07 & High & High & Low & High & High\\
\hline
0.06 & High & High & Low & Low & Low\\
\hline
0.05 & High & Low & Low & High & Low\\
\hline
0.04 & High & High & Low & High & Low\\
\hline
\end{tabular}
\end{table}

\newpage

\textbf{S6 Table. Highest and lowest probability of injury, conditional on the all variables in the Markov blanket for ``injured''.}

\begin{table}[H]
\centering
\begin{tabular}{c|c|c|c|c}
\hline
\textbf{Probability} & \textbf{hours} & \textbf{pi} & \textbf{nle} & \textbf{stiffness}\\
\hline
0.77 & High & injury & +1SD & +1SD\\
\hline
0.74 & High & no injury & +1SD & +1SD\\
\hline
0.72 & Low & injury & +1SD & +1SD\\
\hline
0.70 & High & injury & +1SD & mean\\
\hline
0.69 & High & injury & +1SD & -1SD\\
\hline
0.68 & High & no injury & +1SD & mean\\
\hline
0.66 & High & no injury & +1SD & -1SD\\
\hline
0.65 & Low & injury & +1SD & mean\\
\hline
0.64 & Low & injury & +1SD & -1SD\\
\hline
0.61 & High & injury & mean & +1SD\\
\hline
0.58 & High & no injury & mean & +1SD\\
\hline
0.54 & Low & injury & mean & +1SD\\
\hline
0.54 & High & injury & mean & mean\\
\hline
0.51 & High & injury & -1SD & +1SD\\
\hline
0.51 & High & injury & mean & -1SD\\
\hline
0.50 & High & no injury & mean & mean\\
\hline
0.48 & High & no injury & mean & -1SD\\
\hline
0.48 & Low & injury & mean & mean\\
\hline
0.47 & High & no injury & -1SD & +1SD\\
\hline
0.46 & Low & injury & mean & -1SD\\
\hline
0.44 & Low & injury & -1SD & +1SD\\
\hline
0.44 & High & injury & -1SD & mean\\
\hline
0.42 & High & injury & -1SD & -1SD\\
\hline
0.41 & High & no injury & -1SD & mean\\
\hline
0.38 & High & no injury & -1SD & -1SD\\
\hline
0.37 & Low & no injury & +1SD & +1SD\\
\hline
0.37 & Low & injury & -1SD & mean\\
\hline
0.35 & Low & injury & -1SD & -1SD\\
\hline
0.32 & Low & no injury & +1SD & mean\\
\hline
0.29 & Low & no injury & +1SD & -1SD\\
\hline
0.20 & Low & no injury & mean & +1SD\\
\hline
0.18 & Low & no injury & mean & mean\\
\hline
0.16 & Low & no injury & mean & -1SD\\
\hline
0.15 & Low & no injury & -1SD & +1SD\\
\hline
0.13 & Low & no injury & -1SD & mean\\
\hline
0.11 & Low & no injury & -1SD & -1SD\\
\hline
\end{tabular}
\end{table}

\end{document}